\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-10-27 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
modifs réunions

As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by \textbf{ACTUALISER LES CHIFFRES -2.1\% per year for mean air transport costs and -2.0\% per year XXX JH à GD : STP, refaire tourner la regression sous-jacente.} for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
\cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture.
The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping. Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip

@@

\begin{table}[htbp]
	\centering
	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
	%\input{Table_baseline10.tex}
	\input{Table_baseline10_reshaped.tex}
	\begin{tablenotes}
		\scriptsize
		\item Statistics for 1997-1999 and 2002-2019
		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
        \item Transport costs expressed in \% of the fas price (except for $\widehat{t}$ expressed in USD.
		%\item \textbf{Model (A): With ad-valorem transport costs only}
		\item Model (B): With additive and ad-valorem transport costs
	\end{tablenotes}
\end{table}

@@

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
International transport costs were reduced by \textbf{50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping. AJUSTER LP}. Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (a), (d) and (e)).
Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (A) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing). Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
This is particularly the case for the multiplicative component (panel (d)).\smallskip

>>>>

As shown in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -1.9\% for mean vessel transport costs, implying a 60\% decrease in both air and vessel over the period.\footnote{One may be puzzled by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for vessel transport.
\cite{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates.
In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:Trends_in_TC} springs from high oil-shock-related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture.
The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs.
We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time-trend analysis.} Using US data, \cite{Hummels_1999} found that overall transport cost declined from 6\% to 4\% of the import value between 1974 and 1996.
For the same years, we obtain a total decrease from 5.3\% to 3\% in terms of the export price on average for air shipping, and from 8.9\% to 4.6\% for ocean shipping. Overall, our results display trends confirm and extend those reported by \cite{Hummels_1999} up to the very recent years.\smallskip

@@

\begin{table}[htbp]
	\centering
	\caption{Transport costs estimates: Robustness to aggregation (2005-2019)} \vspace{5mm} \label{tab:baseline10}
	\input{Table_baseline10.tex}
	\input{Table_baseline10_reshaped.tex}
	\begin{tablenotes}
		\scriptsize
		\item Statistics for 1997-1999 and 2002-2019
		\item Statistics are obtained weighting each observation by its value relative to total trade flows.
        \item Transport costs expressed in \% of the fas price (except for $\widehat{t}$ expressed in USD.
		%\item \textbf{Model (A): With ad-valorem transport costs only}
		\item Model (B): With additive and ad-valorem transport costs
	\end{tablenotes}
\end{table}

@@

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}.
First, in accordance with Figure \ref{fig:Trends_in_TC}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)).
International transport costs were reduced by 50\% between 1974 and 2019 in air shipping, and by 60\% in maritime shipping. Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (a), (d) and (e)).
Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs.
Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for air transport (panels (A) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs.
For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period.
Air transport costs were reduced by 50\% between 1974 and 2019, which is mainly attributable to a reduction in the transport costs ``per se''.
Composition effects are stronger in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)): they amplify the reduction in ``pure'' transport costs at the beginning of the period, in the 1970s (the dotted line is below the plain line, and the gap is increasing). Afterwards, the gap between the two remains roughly constant across years, indicating that the amplification effect does not strengthen anymore: the two trends are similar.
Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed into a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two).
This is particularly the case for the multiplicative component (panel (d)).\smallskip
